### 2024-02-18 author-kit-CVPR2024-v2/sec/0_abstract.tex
Formatted header and added abstract for proposal

\begin{abstract}
    Dance is a universal form of expression and entertainment, yet mastering its intricacies
     can be daunting for beginners. To democratize dance learning and make it more accessible
    to enthusiasts, we propose a novel approach utilizing computer vision techniques. By
    breaking down dance videos into vectorized and text-based representations with counts,
    the complexities of dance routines are simplified, enabling easier comprehension and
    replication by average fans.
\end{abstract}