### 2024-04-25 author-kit-CVPR2024-v2/sec/1_intro.tex
updated to PT

\section{Introduction}
\label{sec:intro}

\indent Over 50 million Americans seek physical therapy every year, whether that be from 
musculoskeletal conditions, car accidents, or sports injuries. Evidently, there is a large
need for physical therapy. However, only 35% of PT patients actually adhere to their assigned 
plan. This is a significant problem since the majority of patients are not following 
through with their exercises and thus decrease the effectiveness of their therapy plan.
\\
\indent We aim to use Computer Vision to partition each exercise movement into more easily digestible chunks 
by creating a model that can learn (vectorize) each exercise movement from two given videos, an instructional 
one and a patient one, and compare the two. By developing a tool that provides feedback as a result of analyzing 
the two videos, which ideally should match perfectly, adhering to an exercise plan 
becomes much more feasible, enjoyable, and even motivating.

%-------------------------------------------------------------------------
\subsection{Existing Attempts}
\indent Several existing attempts have been made to tackle the challenge of improving patient adherence, 
each with its own approach and limitations. One such approach involves utilizing monitoring technology, 
such as pedometers, which have proved to be user-friendly. Though such monitoring technologies are 
able to monitor, or track, the patients' physical activity decently well, there was not much of an effect in terms 
of improving their adherence to their exercise plans. As a result, patients using such technologies still struggle with 
adhering to their exercise plans.
\\
\indent Another attempt that has been made is virtual and augmented reality treatment. Through this method, patients 
are able to put on their VR headsets and transport into any environment they would like to complete their exercises in. 
The possibilities are endless: they could surround themselves in peaceful, serene nature, or they could 
opt for a more competitive gaming atmosphere. Physical therapists have found that this actually 
makes their treatment plans more appealing and thus improves their patient adherence rates. However, 
it is important to note that this is significantly more expensive. The cost associated with this 
method is typically too high for the majority of patients and thus makes this approach less practical.
\\
\indent Unlike the aforementioned methods, our proposed method is low-cost, accessible, and effective. 
Implementing the aspect of gamification allows for physical therapy to be fun yet productive. Patients become 
competitors, working hard to beat their previous scores. Additionally, our method could be expanded to offer 
further personalized feedback based on the analysis of the learner's own video and the given instructional video, 
ideally without the requirement for perfect lighting or camera angle conditions.

\subsection{Useful Techniques}
\indent Several computer vision techniques offer promising avenues for addressing the challenge of learning 
exercising movements from instructional videos by providing deeper insights into their dynamics and nuances. 
One such technique involves exploring 3D models instead of traditional 2D skeletonization
 methods to create more comprehensive representations of the exercise movements. Unlike 2D skeletonization, 
 which captures only the spatial coordinates of key body joints in a single plane, 3D modeling 
 techniques can capture the full spatial extent of exericse movements, including depth information\cite{Alpher03}. This 
 allows for a more accurate and detailed representation of the individual's body and movements in 
 three-dimensional space, which is particularly valuable for capturing complex and multi-dimensional 
 exercise movements\cite{Authors14b}.
\\
\indent In addition to 3D modeling, other computer vision techniques such as action recognition, pose 
estimation, trajectory analysis, and video segmentation and tracking can provide valuable insights 
into the dynamics of exercise movements. Action recognition algorithms can automatically identify and 
classify different gestures and even exercise sequences, allowing for automated analysis of exercise 
patterns and styles. Pose estimation techniques can accurately estimate the skeletal poses of 
people from video frames, providing a structural representation of their movements that can be 
used for further analysis and interpretation.
\\
\indent Trajectory analysis techniques can track the trajectories of specific body parts or exercise movements 
over time, allowing for the analysis of exercise movement patterns, velocities, and accelerations\cite{Alpher02}. This 
provides valuable insights into the dynamic aspects of these exercises, 
helping learners to better understand the timing and pacing of the movements. Video segmentation and 
tracking methods also serve to segment the videos into individual sequences or movements, allowing for more 
focused analysis and comparison of specific elements.
\\
\indent By leveraging these advanced computer vision techniques, learners can gain a deeper understanding 
of the dynamics and nuances of exercise movements, facilitating more effective learning experiences. 
These techniques can help learners to break down complex movements into smaller, more 
manageable components, identify areas for improvement, and refine their technique with greater 
precision and accuracy. Overall, the integration of computer vision into physical therapy has the 
potential to revolutionize the way patients approach their recovery plan, making it more accessible, engaging, 
and effective for patients of all levels.

\subsection{Challenges}
\indent Several challenges arise from attempting this project. One is the quality of input data, which 
encompasses various factors such as lighting, angles, occlusions, frame rates, and blurring. These 
factors can significantly impact the accuracy of movement analysis, as they can obscure or distort 
key visual information needed for understanding exercise movements. Addressing these challenges may 
require the development of robust computer vision algorithms capable of handling noisy or low-quality 
video data, as well as techniques for preprocessing and enhancing the visual information to improve 
analysis accuracy.
\\
\indent Moreover, handling interactions between multiple persons in an input video presents additional 
complexities in accurately capturing and representing exercise movements. In some videos, 
there may be two individuals present that interact with each other in complex ways. It is not uncommon for  
one to assist the other in certain movements, for example, by helping lift up or stretch their legs. 
Capturing these interactions and accurately representing them in the analysis poses challenges for 
computer vision algorithms, as they must be able to differentiate between individuals 
and track their movements independently while also accounting for their interactions with others. 
Addressing this challenge may require the development of advanced tracking and segmentation techniques 
capable of accurately identifying and tracking multiple persons in a dynamic environment.
\\
\indent Overall, addressing these challenges requires a multidisciplinary approach that combines expertise 
in computer vision, exercise analysis, and human-computer interaction. By developing robust algorithms, 
incorporating domain-specific knowledge, and leveraging advanced techniques for handling complex 
visual data, it is possible to create a tool that effectively supports learning exercises from  
videos using computer vision. However, continued research and development efforts are needed to 
overcome these challenges and realize the full potential of computer vision technology in the 
field of physical therapy.

\subsection{Technical Gap}
\indent The primary aim of this project is to reduce the technical gap between physical therapy videos and 
effective recovery plans. By enabling the correspondence between individuals with different body proportions 
captured from various angles, this tool seeks to enhance the accessibility and inclusivity of these exercises plans. 
This interdisciplinary endeavor involves collaboration between experts in physical therapy, biomechanics, and computer 
vision, with computer vision providing quantitative analysis capabilities to enhance the understanding and 
teaching of exercise movements.
\\
\indent In conclusion, developing a tool for learning exercises from physical therapy videos using computer vision holds 
great potential for revolutionizing physical therapy plans and making them more accessible and doable to a wider 
audience. By leveraging advanced computer vision techniques and interdisciplinary collaboration, 
this project aims to bridge the gap between physical therapy videos and effective recovery plans, 
ultimately empowering individuals to learn and appreciate physical therapy in new and innovative ways.

% \subsection{Paper length}
% Papers, excluding the references section, must be no longer than eight pages in length.
% The references section will not be included in the page count, and there is no limit on the length of the references section.
% For example, a paper of eight pages with two pages of references would have a total length of 10 pages.
% {\bf There will be no extra page charges for \confName\ \confYear.}

% Overlength papers will simply not be reviewed.
% This includes papers where the margins and formatting are deemed to have been significantly altered from those laid down by this style guide.
% Note that this \LaTeX\ guide already sets figure captions and references in a smaller font.
% The reason such papers will not be reviewed is that there is no provision for supervised revisions of manuscripts.
% The reviewing process cannot determine the suitability of the paper for presentation in eight pages if it is reviewed in eleven.

% %-------------------------------------------------------------------------
% \subsection{The ruler}
% The \LaTeX\ style defines a printed ruler which should be present in the version submitted for review.
% The ruler is provided in order that reviewers may comment on particular lines in the paper without circumlocution.
% If you are preparing a document using a non-\LaTeX\ document preparation system, please arrange for an equivalent ruler to appear on the final output pages.
% The presence or absence of the ruler should not change the appearance of any other content on the page.
% The camera-ready copy should not contain a ruler.
% (\LaTeX\ users may use options of \texttt{cvpr.sty} to switch between different versions.)

% Reviewers:
% note that the ruler measurements do not align well with lines in the paper --- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
% Just use fractional references (\eg, this line is $087.5$), although in most cases one would expect that the approximate location will be adequate.

% \subsection{Paper ID}
% Make sure that the Paper ID from the submission system is visible in the version submitted for review (replacing the ``*****'' you see in this document).
% If you are using the \LaTeX\ template, \textbf{make sure to update paper ID in the appropriate place in the tex file}.

% \subsection{Mathematics}

% Please number all of your sections and displayed equations as in these examples:
% \begin{equation}
%   E = m\cdot c^2
%   \label{eq:important}
% \end{equation}
% and
% \begin{equation}
%   v = a\cdot t.
%   \label{eq:also-important}
% \end{equation}
% It is important for readers to be able to refer to any particular equation.
% Just because you did not refer to it in the text does not mean some future reader might not need to refer to it.
% It is cumbersome to have to use circumlocutions like ``the equation second from the top of page 3 column 1''.
% (Note that the ruler will not be present in the final copy, so is not an alternative to equation numbers).
% All authors will benefit from reading Mermin's description of how to write mathematics:
% \url{http://www.pamitc.org/documents/mermin.pdf}.

% \subsection{Blind review}

% Many authors misunderstand the concept of anonymizing for blind review.
% Blind review does not mean that one must remove citations to one's own work---in fact it is often impossible to review a paper unless the previous citations are known and available.

% Blind review means that you do not use the words ``my'' or ``our'' when citing previous work.
% That is all.
% (But see below for tech reports.)

% Saying ``this builds on the work of Lucy Smith [1]'' does not say that you are Lucy Smith;
% it says that you are building on her work.
% If you are Smith and Jones, do not say ``as we show in [7]'', say ``as Smith and Jones show in [7]'' and at the end of the paper, include reference 7 as you would any other cited work.

% An example of a bad paper just asking to be rejected:
% \begin{quote}
% \begin{center}
%     An analysis of the frobnicatable foo filter.
% \end{center}

%    In this paper we present a performance analysis of our previous paper [1], and show it to be inferior to all previously known methods.
%    Why the previous paper was accepted without this analysis is beyond me.

%    [1] Removed for blind review
% \end{quote}

% An example of an acceptable paper:
% \begin{quote}
% \begin{center}
%      An analysis of the frobnicatable foo filter.
% \end{center}

%    In this paper we present a performance analysis of the  paper of Smith \etal [1], and show it to be inferior to all previously known methods.
%    Why the previous paper was accepted without this analysis is beyond me.

%    [1] Smith, L and Jones, C. ``The frobnicatable foo filter, a fundamental contribution to human knowledge''. Nature 381(12), 1-213.
% \end{quote}

% If you are making a submission to another conference at the same time, which covers similar or overlapping material, you may need to refer to that submission in order to explain the differences, just as you would if you had previously published related work.
% In such cases, include the anonymized parallel submission~\cite{Authors14} as supplemental material and cite it as
% \begin{quote}
% [1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324, Supplied as supplemental material {\tt fg324.pdf}.
% \end{quote}

% Finally, you may feel you need to tell the reader that more details can be found elsewhere, and refer them to a technical report.
% For conference submissions, the paper must stand on its own, and not {\em require} the reviewer to go to a tech report for further details.
% Thus, you may say in the body of the paper ``further details may be found in~\cite{Authors14b}''.
% Then submit the tech report as supplemental material.
% Again, you may not assume the reviewers will read this material.

% Sometimes your paper is about a problem which you tested using a tool that is widely known to be restricted to a single institution.
% For example, let's say it's 1969, you have solved a key problem on the Apollo lander, and you believe that the 1970 audience would like to hear about your
% solution.
% The work is a development of your celebrated 1968 paper entitled ``Zero-g frobnication: How being the only people in the world with access to the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

% You can handle this paper like any other.
% Do not write ``We show how to improve our previous work [Anonymous, 1968].
% This time we tested the algorithm on a lunar lander [name of lander removed for blind review]''.
% That would be silly, and would immediately identify the authors.
% Instead write the following:
% \begin{quotation}
% \noindent
%    We describe a system for zero-g frobnication.
%    This system is new because it handles the following cases:
%    A, B.  Previous systems [Zeus et al. 1968] did not  handle case B properly.
%    Ours handles it by including a foo term in the bar integral.

%    ...

%    The proposed system was integrated with the Apollo lunar lander, and went all the way to the moon, don't you know.
%    It displayed the following behaviours, which show how well we solved cases A and B: ...
% \end{quotation}
% As you can see, the above text follows standard scientific convention, reads better than the first version, and does not explicitly name you as the authors.
% A reviewer might think it likely that the new paper was written by Zeus \etal, but cannot make any decision based on that guess.
% He or she would have to be sure that no other authors could have been contracted to solve problem B.
% \medskip

% \noindent
% FAQ\medskip\\
% {\bf Q:} Are acknowledgements OK?\\
% {\bf A:} No.  Leave them for the final copy.\medskip\\
% {\bf Q:} How do I cite my results reported in open challenges?
% {\bf A:} To conform with the double-blind review policy, you can report results of other challenge participants together with your results in your paper.
% For your results, however, you should not identify yourself and should not mention your participation in the challenge.
% Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\

% \begin{figure}[t]
%   \centering
%   \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
%    %\includegraphics[width=0.8\linewidth]{egfigure.eps}

%    \caption{Example of caption.
%    It is set in Roman so that mathematics (always set in Roman: $B \sin A = A \sin B$) may be included without an ugly clash.}
%    \label{fig:onecol}
% \end{figure}

% \subsection{Miscellaneous}

% \noindent
% Compare the following:\\
% \begin{tabular}{ll}
%  \verb'$conf_a$' &  $conf_a$ \\
%  \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
% \end{tabular}\\
% See The \TeX book, p165.

% The space after \eg, meaning ``for example'', should not be a sentence-ending space.
% So \eg is correct, {\em e.g.} is not.
% The provided \verb'\eg' macro takes care of this.

% When citing a multi-author paper, you may save space by using ``et alia'', shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word).
% If you use the \verb'\etal' macro provided, then you need not worry about double periods when used at the end of a sentence as in Alpher \etal.
% However, use it only when there are three or more authors.
% Thus, the following is correct:
%    ``Frobnication has been trendy lately.
%    It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
%    Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

% This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...'' because reference~\cite{Alpher03} has just two authors.

% \begin{figure*}
%   \centering
%   \begin{subfigure}{0.68\linewidth}
%     \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
%     \caption{An example of a subfigure.}
%     \label{fig:short-a}
%   \end{subfigure}
%   \hfill
%   \begin{subfigure}{0.28\linewidth}
%     \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
%     \caption{Another example of a subfigure.}
%     \label{fig:short-b}
%   \end{subfigure}
%   \caption{Example of a short caption, which should be centered.}
%   \label{fig:short}
% \end{figure*}