### 2024-05-04 paper/sec/0_abstract.tex
add files and restructure directories in anticipation of code review

\begin{abstract}
Physical therapy can be likened to a 'rite of passage' that many people go through
because of how common it is. It provides a solution to the patient's problem, whether
that be back pain or a broken leg, but patients must be able to complete their assigned
recovery plans in order to fully heal. To optimize patient recovery, we propose a novel
approach to navigating their recovery plans utilizing computer vision techniques. By
breaking down instruction videos into vectorized representations, we can break down the
patient's videos similarly and compare the two, providing feedback and showing the
patient's progress over time, thus providing motivation in their journey.
\end{abstract}